---
title: "Research Project"
author: "Jennifer Habicher"
date: "2024-11-28"
output: "github_document"
---

# load packages and dataset
```{r}
library (haven)
library (plyr)
library (dplyr)
library (ggplot2)
library (psych)
library (car)
library (multcomp)
library (Rmisc)
library (bruceR)

#load Dataset
load("ICPSR_38417/DS0001/38417-0001-Data.rda")

```

```{r}
Selected_Dataset <- da38417.0001 %>%
  dplyr::select(Q17_1, Q19, Q20, Q21, Q22, Q117, Q65, Q55E, D2, HHR5, IDENTITY_1_R)
```

# Recode Variables
```{r}
#Q17_1: Please indicate the level of happiness, all things considered, in your relationship 

#Recoding Extremely Happy and Perfect to both read 6 to keep the composite variable the same
Selected_Dataset$Q17_1 <- dplyr::recode(Selected_Dataset$Q17_1, "(1) Extremely unhappy" = 1, "(2) Fairly unhappy" = 2, "(3) A little unhappy" = 3, "(4) Happy" = 4, "(5) Very happy" = 5, "(6) Extremely happy" = 6, "(7) Perfect" = 6)

#Q19: I have a warm and comfortable relationship with my spouse/partner
Selected_Dataset$Q19 <- dplyr::recode(Selected_Dataset$Q19, "(1) Not at all true" = 1, "(2) A little true" = 2, "(3) Somewhat true" = 3, "(4) Mostly true" = 4, "(5) Almost completely true" = 5, "(6) Completely true" = 6)

#Q20: How rewarding is your relationship with your spouse/partner? 
Selected_Dataset$Q20 <- dplyr::recode(Selected_Dataset$Q20, "(1) Not at all rewarding" = 1, "(2) A little rewarding" = 2, "(3) Somewhat rewarding" = 3, "(4) Mostly rewarding" = 4, "(5) Almost completely rewarding" = 5, "(6) Completely rewarding" = 6)

#Q21: In general, how satisfied are you with your relationship? 
Selected_Dataset$Q21 <- dplyr::recode(Selected_Dataset$Q21, "(1) Not at all satisfied" = 1, "(2) A little satisfied" = 2, "(3) Somewhat satisfied" = 3, "(4) Mostly satisfied" = 4, "(5) Almost completely satisfied" = 5, "(6) Completely satisfied" = 6)

#Q22: In general, how committed are you to your current spouse/partner?
Selected_Dataset$COMMITMENT <- dplyr::recode(Selected_Dataset$Q22, "(1) Not at all committed" = 1, "(2) A little committed" = 2, "(3) Somewhat committed" = 3, "(4) Mostly committed" = 4, "(5) Almost completely committed" = 5, "(6) Completely committed" = 6)

#Q117: How satisfied are you with your sexual relationship with your current spouse/partner? 
Selected_Dataset$SEXUAL_SATISFACTION <- dplyr::recode(Selected_Dataset$Q117, "(1) Very dissatisfied" = 1, "(2) Somewhat dissatisfied" = 2, "(3) Neither satisfied or dissatisfied" = 3, "(4) Somewhat satisfied" = 4, "(5) Very satisfied" = 5)

#Q65: In the past 30 days have you used antidepressant medication
Selected_Dataset$ANTIDEPRESSANT_USE <- droplevels(dplyr::recode(Selected_Dataset$Q65, "(1) Yes" = "Yes", "(2) No" = "No"))

#Q55E: I felt depressed
Selected_Dataset$DEPRESSION_LEVEL <- droplevels(dplyr::recode(Selected_Dataset$Q55E, "(1) Rarely or none of the time (Less than 1 day)" = "Rarely/None", "(2) Some or a little of the time (1-2 days)" = "Some", "(3) Occasionally or a moderate amount of time (3-4 days)" = "Occasionally", "(4) Most or all of the time (5-7 days)" = "Most/All"))

#D2: Which of the following best describes your gender?
Selected_Dataset$GENDER <- droplevels(dplyr::recode(Selected_Dataset$D2, "(1) Man" = "Men", "(2) Woman" = "Women", "(3) Transgender" = "Transgender","(5) Do not identify as any of the above (there is an option to specify at next question)" = "Other"))

#IDENTITY_1_R IDENTITY. Which of the following do you consider yourself to be? (select all that apply) 1. Heterosexual or "straight"
Selected_Dataset$SEXUAL_IDENTITY <- droplevels(dplyr::recode(Selected_Dataset$IDENTITY_1_R, "(1) Heterosexual" = "Heterosexual", "(-99) Valid Non-Response" = "Other"))


#Remove NA values
Selected_Dataset <- na.omit(Selected_Dataset)

```

#Graphs for Recoded Variables
```{r}

#Q17_1: Please indicate the level of happiness, all things considered, in your relationship 
#Grouped by GENDER
ggplot(Selected_Dataset, aes(x = Q17_1)) + facet_wrap(~GENDER) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Relationship Happiness", x = "Happiness", y = "Frequency")

#Grouped by ANTIDEPRESSANT_USE
ggplot(Selected_Dataset, aes(x = Q17_1)) + facet_wrap(~ANTIDEPRESSANT_USE) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Relationship Happiness", x = "Happiness", y = "Frequency")

#Q19: I have a warm and comfortable relationship with my spouse/partner
#Grouped by GENDER
ggplot(Selected_Dataset, aes(x = Q19)) + facet_wrap(~GENDER) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Warm and Comfortable Relationship", x = "Warmth", y = "Frequency")

#Grouped by ANTIDEPRESSANT_USE
ggplot(Selected_Dataset, aes(x = Q19)) + facet_wrap(~ANTIDEPRESSANT_USE) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Warm and Comfortable Relationship", x = "Warmth", y = "Frequency")

#Q20: How rewarding is your relationship with your spouse/partner? 
#Grouped by GENDER
ggplot(Selected_Dataset, aes(x = Q20))  + facet_wrap(~GENDER) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Rewarding Relationship", x = "Reward", y = "Frequency")

#Grouped by ANTIDEPRESSANT_USE
ggplot(Selected_Dataset, aes(x = Q20)) + facet_wrap(~ANTIDEPRESSANT_USE) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Rewarding Relationship", x = "Reward", y = "Frequency")

#Q21: In general, how satisfied are you with your relationship?
#Grouped by GENDER
ggplot(Selected_Dataset, aes(x = Q21))  + facet_wrap(~GENDER) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Relationship Satisfaction", x = "Satisfaction", y = "Frequency")

#Grouped by ANTIDEPRESSANT_USE
ggplot(Selected_Dataset, aes(x = Q21)) + facet_wrap(~ANTIDEPRESSANT_USE) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Relationship Satisfaction", x = "Satisfaction", y = "Frequency")

#Q22: In general, how committed are you to your current spouse/partner? (Numeric)
#Grouped by GENDER
ggplot(Selected_Dataset, aes(x = COMMITMENT))  + facet_wrap(~GENDER) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Relationship Commitment", x = "Commitment", y = "Frequency")

#Grouped by ANTIDEPRESSANT_USE
ggplot(Selected_Dataset, aes(x = COMMITMENT)) + facet_wrap(~ANTIDEPRESSANT_USE) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Relationship Commitment", x = "Commitment", y = "Frequency")

#Q117: How satisfied are you with your sexual relationship with your current spouse/partner? 
#Grouped by GENDER
ggplot(Selected_Dataset, aes(x = SEXUAL_SATISFACTION))  + facet_wrap(~GENDER) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Sexual Relationship Satisfaction", x = "Satisfaction", y = "Frequency")

#Grouped by ANTIDEPRESSANT_USE
ggplot(Selected_Dataset, aes(x = SEXUAL_SATISFACTION)) + facet_wrap(~ANTIDEPRESSANT_USE) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Sexual Relationship Satisfaction", x = "Satisfaction", y = "Frequency")


```


# Create Composite Variable
```{r}
#For Relationship Satisfaction I will produce one composite variable out of variables Q17_1, Q19, Q20, Q21
#Mutate and create a new column. Use the average of each participants scores
Selected_Dataset$Relationship_Satisfaction <- rowMeans(Selected_Dataset %>% dplyr::select(Q17_1, Q19, Q20, Q21), na.rm = TRUE)

#plot the composite variable
ggplot(Selected_Dataset, aes(x = Relationship_Satisfaction)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Relationship Satisfaction Composite Variable", x = "Relationship Satisfaction", y = "Frequency")

#Group By GENDER
ggplot(Selected_Dataset, aes(x = Relationship_Satisfaction)) + facet_wrap(~GENDER) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Relationship Satisfaction Composite Variable", x = "Relationship Satisfaction", y = "Frequency")

#Group By ANTIDEPRESSANT_USE
ggplot(Selected_Dataset, aes(x = Relationship_Satisfaction)) + facet_wrap(~ANTIDEPRESSANT_USE) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Relationship Satisfaction Composite Variable", x = "Relationship Satisfaction", y = "Frequency")

```



#Check Normality of Distribution for each variable

```{r}

kruskal.test(Selected_Dataset$Relationship_Satisfaction ~ Selected_Dataset$GENDER)
kruskal.test(Selected_Dataset$Relationship_Satisfaction ~ Selected_Dataset$ANTIDEPRESSANT_USE)

kruskal.test(Selected_Dataset$SEXUAL_SATISFACTION ~ Selected_Dataset$GENDER)
kruskal.test(Selected_Dataset$SEXUAL_SATISFACTION ~ Selected_Dataset$ANTIDEPRESSANT_USE)

kruskal.test(Selected_Dataset$COMMITMENT ~ Selected_Dataset$GENDER)
kruskal.test(Selected_Dataset$COMMITMENT ~ Selected_Dataset$ANTIDEPRESSANT_USE)

```

#Secondary Normality Check
```{r}
table(Selected_Dataset$GENDER, Selected_Dataset$ANTIDEPRESSANT_USE)

```
Since the tests for Normality has failed for Gender, I will merge Transgenger and Other to a single group called Other

```{r}
# Merge Transgender and Other into a single group called Other
Selected_Dataset$GENDER <- droplevels(dplyr::recode(Selected_Dataset$GENDER, "Transgender" = "Other"))

table(Selected_Dataset$GENDER, Selected_Dataset$ANTIDEPRESSANT_USE)

```

Verified that now all groups have enough N to satisfy Central Limit Theorem. 


#Make Graphs for each Variable
```{r}
#Plot
ggplot(Selected_Dataset, aes(x = ANTIDEPRESSANT_USE, y = Relationship_Satisfaction, fill = GENDER)) +
  geom_boxplot(outlier.shape = 16, outlier.colour = "black", outlier.size = 1) + 
  labs(
    title = "Relationship Satisfaction by Antidepressant Use",
    x = "Antidepressant Use",
    y = "Relationship Satisfaction"
  ) +
  theme_linedraw(base_size = 20) + 
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18, hjust = 1), 
    axis.text.y = element_text(size = 18),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 15),
    legend.position = "top",
    legend.margin = margin(t = -5, b = -5), # Adjust top and bottom margin
    legend.spacing.y = unit(0, "pt")       # Reduce vertical spacing between legend items

  ) +
  coord_cartesian(clip = 'off') + # Avoid clipping of axis labels if needed
  expand_limits(y = c(min(Selected_Dataset$Relationship_Satisfaction), max(Selected_Dataset$Relationship_Satisfaction))) + # Ensures y-axis limits are fully shown
  scale_y_continuous(
    breaks = seq(min(Selected_Dataset$Relationship_Satisfaction), 
                 max(Selected_Dataset$Relationship_Satisfaction), 
                 by = 1) # Adjust the step size as needed
  ) + scale_fill_manual( values = c("Men" = "red", "Women" = "orange", "Other"="blue") )


#Sexual Satisfaction
ggplot(Selected_Dataset, aes(x = ANTIDEPRESSANT_USE, y = SEXUAL_SATISFACTION, fill = GENDER)) + 
  geom_boxplot(outlier.shape = 16, outlier.colour = "black", outlier.size = 1) + 
  labs(
    title = "Sexual Satisfaction by Antidepressant Use",
    x = "Antidepressant Use",
    y = "Sexual Satisfaction",
  ) +
  theme_linedraw(base_size = 20) +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18, hjust = 1), 
    axis.text.y = element_text(size = 18),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 15),
    legend.position = "top",
    legend.margin = margin(t = -5, b = -5), # Adjust top and bottom margin
    legend.spacing.y = unit(0, "pt")       # Reduce vertical spacing between legend items
  ) +
  coord_cartesian(clip = 'off') + # Avoid clipping of axis labels if needed
  expand_limits(y = c(1, 6)) + # Ensures y-axis limits are fully shown
  scale_y_continuous(
    breaks = seq(1, 6, by = 1) # Adjust the step size as needed
  )+ scale_fill_manual( values = c("Men" = "red", "Women" = "orange", "Other"="blue") )

# Commitment

ggplot(Selected_Dataset, aes(x = ANTIDEPRESSANT_USE, y = COMMITMENT, fill = GENDER)) + 
  geom_boxplot(outlier.shape = 16, outlier.colour = "black", outlier.size = 1) + 
  labs(
    title = "Commitment by Antidepressant Use",
    x = "Antidepressant Use",
    y = "Commitment",
  ) +
  theme_linedraw(base_size = 20) +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18, hjust = 1), 
    axis.text.y = element_text(size = 18),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 17),
    legend.position = "top",
    legend.margin = margin(t = -5, b = -5), # Adjust top and bottom margin
    legend.spacing.y = unit(0, "pt")       # Reduce vertical spacing between legend items
  ) +
  coord_cartesian(clip = 'off') + # Avoid clipping of axis labels if needed
  expand_limits(y = c(min(Selected_Dataset$COMMITMENT), max(Selected_Dataset$COMMITMENT))) + # Ensures y-axis limits are fully shown
  scale_y_continuous(
    breaks = seq(min(Selected_Dataset$COMMITMENT), 
                 max(Selected_Dataset$COMMITMENT), 
                 by = 1) # Adjust the step size as needed
  )+ scale_fill_manual( values = c("Men" = "red", "Women" = "orange", "Other"="blue") )

```

#Make Graphs for each Variable wihtout other
```{r}
#Plot
# Filter the dataset to include only "Men" and "Women"
Filtered_Dataset <- Selected_Dataset %>% 
  filter(GENDER %in% c("Men", "Women"))

# Plot
ggplot(Filtered_Dataset, aes(x = ANTIDEPRESSANT_USE, y = Relationship_Satisfaction, fill = GENDER)) +
  geom_boxplot(outlier.shape = 16, outlier.colour = "black", outlier.size = 1) + 
  labs(
    title = "Relationship Satisfaction by Antidepressant Use",
    x = "Antidepressant Use",
    y = "Relationship Satisfaction"
  ) +
  theme_linedraw(base_size = 20) + 
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18, hjust = 1), 
    axis.text.y = element_text(size = 18),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 15),
    legend.position = "top",
    legend.margin = margin(t = -5, b = -5), # Adjust top and bottom margin
    legend.spacing.y = unit(0, "pt")       # Reduce vertical spacing between legend items
  ) +
  coord_cartesian(clip = 'off') + # Avoid clipping of axis labels if needed
  expand_limits(y = c(min(Filtered_Dataset$Relationship_Satisfaction), max(Filtered_Dataset$Relationship_Satisfaction))) + # Ensures y-axis limits are fully shown
  scale_y_continuous(
    breaks = seq(min(Filtered_Dataset$Relationship_Satisfaction), 
                 max(Filtered_Dataset$Relationship_Satisfaction), 
                 by = 1) # Adjust the step size as needed
  ) + 
  scale_fill_manual(values = c("Men" = "red", "Women" = "orange")) # Remove the color for "Other"


#Sexual Satisfaction
ggplot(Filtered_Dataset, aes(x = ANTIDEPRESSANT_USE, y = SEXUAL_SATISFACTION, fill = GENDER)) + 
  geom_boxplot(outlier.shape = 16, outlier.colour = "black", outlier.size = 1) + 
  labs(
    title = "Sexual Satisfaction by Antidepressant Use",
    x = "Antidepressant Use",
    y = "Sexual Satisfaction",
  ) +
  theme_linedraw(base_size = 20) +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18, hjust = 1), 
    axis.text.y = element_text(size = 18),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 15),
    legend.position = "top",
    legend.margin = margin(t = -5, b = -5), # Adjust top and bottom margin
    legend.spacing.y = unit(0, "pt")       # Reduce vertical spacing between legend items
  ) +
  coord_cartesian(clip = 'off') + # Avoid clipping of axis labels if needed
  expand_limits(y = c(1, 6)) + # Ensures y-axis limits are fully shown
  scale_y_continuous(
    breaks = seq(1, 6, by = 1) # Adjust the step size as needed
  )+ scale_fill_manual( values = c("Men" = "red", "Women" = "orange") )

# Commitment

ggplot(Filtered_Dataset, aes(x = ANTIDEPRESSANT_USE, y = COMMITMENT, fill = GENDER)) + 
  geom_boxplot(outlier.shape = 16, outlier.colour = "black", outlier.size = 1) + 
  labs(
    title = "Commitment by Antidepressant Use",
    x = "Antidepressant Use",
    y = "Commitment",
  ) +
  theme_linedraw(base_size = 20) +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18, hjust = 1), 
    axis.text.y = element_text(size = 18),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 17),
    legend.position = "top",
    legend.margin = margin(t = -5, b = -5), # Adjust top and bottom margin
    legend.spacing.y = unit(0, "pt")       # Reduce vertical spacing between legend items
  ) +
  coord_cartesian(clip = 'off') + # Avoid clipping of axis labels if needed
  expand_limits(y = c(min(Filtered_Dataset$COMMITMENT), max(Filtered_Dataset$COMMITMENT))) + # Ensures y-axis limits are fully shown
  scale_y_continuous(
    breaks = seq(min(Filtered_Dataset$COMMITMENT), 
                 max(Filtered_Dataset$COMMITMENT), 
                 by = 1) # Adjust the step size as needed
  )+ scale_fill_manual( values = c("Men" = "red", "Women" = "orange") )


# Filter the dataset to include only "Men" and "Women"
Filtered_Dataset <- Selected_Dataset %>% 
  filter(GENDER %in% c("Men", "Women"))

# Plot density curves
ggplot(Filtered_Dataset, aes(x = Relationship_Satisfaction, color = interaction(GENDER, ANTIDEPRESSANT_USE), fill = interaction(GENDER, ANTIDEPRESSANT_USE))) +
  geom_density(alpha = 0.4) +  # Transparency for overlapping areas
  labs(
    title = "Density Plot of Relationship Satisfaction by Gender and Antidepressant Use",
    x = "Relationship Satisfaction",
    y = "Density"
  ) +
  theme_linedraw(base_size = 20) +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18), 
    axis.text.y = element_text(size = 18),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 15),
    legend.position = "top",
    legend.margin = margin(t = -5, b = -5), # Adjust top and bottom margin
    legend.spacing.y = unit(0, "pt")       # Reduce vertical spacing between legend items
  ) +
  scale_color_manual(
    values = c(
      "Men.FALSE" = "blue", "Men.TRUE" = "darkblue",
      "Women.FALSE" = "orange", "Women.TRUE" = "darkorange"
    ),
    labels = c(
      "Men: No Antidepressant", "Men: Antidepressant",
      "Women: No Antidepressant", "Women: Antidepressant"
    )
  ) +
  scale_fill_manual(
    values = c(
      "Men.FALSE" = "blue", "Men.TRUE" = "darkblue",
      "Women.FALSE" = "orange", "Women.TRUE" = "darkorange"
    ),
    labels = c(
      "Men: No Antidepressant", "Men: Antidepressant",
      "Women: No Antidepressant", "Women: Antidepressant"
    )
  )


# Filter the dataset to include only "Men" and "Women"
Filtered_Dataset <- Selected_Dataset %>% 
  filter(GENDER %in% c("Men", "Women"))

# Plot density curves as lines
ggplot(Filtered_Dataset, aes(x = Relationship_Satisfaction, color = interaction(GENDER, ANTIDEPRESSANT_USE))) +
  geom_density(size = 1.2) +  # Line size for better visibility
  labs(
    title = "Density Plot of Relationship Satisfaction by Gender and Antidepressant Use",
    x = "Relationship Satisfaction",
    y = "Density"
  ) +
  theme_linedraw(base_size = 20) +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18), 
    axis.text.y = element_text(size = 18),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 15),
    legend.position = "top",
    legend.margin = margin(t = -5, b = -5), # Adjust top and bottom margin
    legend.spacing.y = unit(0, "pt")       # Reduce vertical spacing between legend items
  ) +
  scale_color_manual(
    values = c(
      "Men.FALSE" = "blue", "Men.TRUE" = "darkblue",
      "Women.FALSE" = "orange", "Women.TRUE" = "darkorange"
    ),
    labels = c(
      "Men: No Antidepressant", "Men: Antidepressant",
      "Women: No Antidepressant", "Women: Antidepressant"
    )
  )



```


#Check for Equality of Variance

```{r}
#Levine's Test

leveneTest(Selected_Dataset$Relationship_Satisfaction ~ Selected_Dataset$GENDER)
# Fail to reject Null Hypothesis: Variances are equal

leveneTest(Selected_Dataset$Relationship_Satisfaction ~ Selected_Dataset$ANTIDEPRESSANT_USE)
# Fail to reject Null Hypothesis: Variances are equal


leveneTest(Selected_Dataset$SEXUAL_SATISFACTION ~ Selected_Dataset$GENDER)
# Reject Null Hypothesis: Variances are not equal -- Use Welch ANOVA 

leveneTest(Selected_Dataset$SEXUAL_SATISFACTION ~ Selected_Dataset$ANTIDEPRESSANT_USE)
# Fail to reject Null Hypothesis: Variances are equal


leveneTest(Selected_Dataset$COMMITMENT ~ Selected_Dataset$GENDER)
# Reject Null Hypothesis: Variances are not equal -- Use Welch ANOVA 

leveneTest(Selected_Dataset$COMMITMENT ~ Selected_Dataset$ANTIDEPRESSANT_USE)
# Fail to reject Null Hypothesis: Variances are equal

```

Sexual Satisfaction and Commitment do not pass the assumption of homogeneity of variance. I will do a log transformation of the data

```{r}
Selected_Dataset$SEXUAL_SATISFACTION <- log(Selected_Dataset$SEXUAL_SATISFACTION)
Selected_Dataset$COMMITMENT <- log(Selected_Dataset$COMMITMENT)
```

Recheck for Equality of Variance

```{r}
leveneTest(Selected_Dataset$SEXUAL_SATISFACTION ~ Selected_Dataset$GENDER)

leveneTest(Selected_Dataset$COMMITMENT ~ Selected_Dataset$GENDER)

```

```{r}
#Summarize Data
summary(Selected_Dataset)
```

#Check Independent Observation Assumption

Given the nature of the survey, I assume that they design of this survey means that respondents are all independent of each other.


#Violated Assumptions and Fixes

For the tests that do not pass the assumptions of normality, I merged the Trangsenger and Other groups into a single group called Other. This new group has a sufficiently large N to satisfy the Central Limit Theorem.
For the tests that do not pass the assumptions of homogeneity of variance, I have transposed the data using a log transformation. 

#Data Analysis

2x3 Design
- IV: Antidepressants vs No Antidepressants
- IV: Gender: Man vs Woman vs Other

- DV: Relationship Satisfaction
- DV: Sexual Satisfaction
- DV: Commitment

```{r}

#Relationship Satisfaction

#descriptive Statistics
group_by(Selected_Dataset, GENDER, ANTIDEPRESSANT_USE) %>%
  dplyr::summarise(
    mean = mean(Relationship_Satisfaction, na.rm = TRUE),
    sd = sd(Relationship_Satisfaction, na.rm = TRUE)
  )

#2x3 MANOVA
Satisfaction_Model <- aov(Relationship_Satisfaction ~  GENDER*ANTIDEPRESSANT_USE, data = Selected_Dataset)

summary(Satisfaction_Model)

# Post Hoc
TukeyHSD(Satisfaction_Model, which = "GENDER:ANTIDEPRESSANT_USE")
pairwise.t.test(Selected_Dataset$Relationship_Satisfaction, 
                interaction(Selected_Dataset$GENDER, Selected_Dataset$ANTIDEPRESSANT_USE),
                p.adjust.method = "bonferroni")


# Create the violin plot
ggplot(Selected_Dataset, aes(x = interaction(GENDER, ANTIDEPRESSANT_USE), 
                              y = Relationship_Satisfaction, 
                              fill = ANTIDEPRESSANT_USE)) +
  geom_violin(trim = FALSE, alpha = 0.7) +  # Violin plot
  geom_boxplot(width = 0.1, position = position_dodge(0.9), outlier.shape = NA) +  # Add boxplot inside
  stat_summary(fun = mean, geom = "point", shape = 23, size = 3, fill = "white") +  # Add mean points
  labs(
    title = "Distribution of Relationship Satisfaction by Gender and Antidepressant Use",
    x = "Gender and Antidepressant Use",
    y = "Relationship Satisfaction"
  ) +
  scale_fill_manual(values = c("#FF9999", "#99CCFF")) +  # Customize fill colors
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.title = element_blank()
  )




# Calculate summary statistics
summary_stats <- Selected_Dataset %>%
  group_by(GENDER, ANTIDEPRESSANT_USE) %>%
  summarise(
    mean_satisfaction = mean(Relationship_Satisfaction, na.rm = TRUE),
    sd_satisfaction = sd(Relationship_Satisfaction, na.rm = TRUE),
    n = n()
  ) %>%
  mutate(
    se = sd_satisfaction / sqrt(n),  # Standard Error
    ci = qt(0.975, df = n - 1) * se  # 95% Confidence Interval
  )

# Create the bar graph
ggplot(summary_stats, aes(x = interaction(GENDER, ANTIDEPRESSANT_USE), 
                          y = mean_satisfaction, 
                          fill = ANTIDEPRESSANT_USE)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), alpha = 0.8) +
  geom_errorbar(aes(ymin = mean_satisfaction - se, ymax = mean_satisfaction + se), 
                width = 0.2, position = position_dodge(width = 0.9)) +
  labs(
    title = "Mean Relationship Satisfaction by Gender and Antidepressant Use",
    x = "Gender and Antidepressant Use",
    y = "Mean Relationship Satisfaction"
  ) +
  scale_fill_manual(values = c("#FF9999", "#99CCFF")) +  # Customize bar colors
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.title = element_blank()
  )


ggplot(summary_stats, aes(x = interaction(GENDER, ANTIDEPRESSANT_USE), 
                          y = mean_satisfaction, 
                          color = ANTIDEPRESSANT_USE)) +
  geom_point(size = 4, position = position_dodge(width = 0.5)) +  # Mean points
  geom_errorbar(aes(ymin = mean_satisfaction - se, ymax = mean_satisfaction + se), 
                width = 0.2, position = position_dodge(width = 0.5)) +
  labs(
    title = "Mean Relationship Satisfaction by Gender and Antidepressant Use",
    x = "Gender and Antidepressant Use",
    y = "Mean Relationship Satisfaction"
  ) +
  scale_color_manual(values = c("#FF9999", "#99CCFF")) +  # Customize colors
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.title = element_blank()
  )


# Create the box plot with jittered points
ggplot(Selected_Dataset, aes(x = interaction(GENDER, ANTIDEPRESSANT_USE), 
                              y = Relationship_Satisfaction, 
                              fill = ANTIDEPRESSANT_USE)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.5, position = position_dodge(width = 0.9)) +  # Boxplot
  geom_jitter(aes(color = ANTIDEPRESSANT_USE), 
              width = 0.2, alpha = 0.5, size = 1.5) +  # Jittered points
  labs(
    title = "Relationship Satisfaction by Gender and Antidepressant Use",
    x = "Gender and Antidepressant Use",
    y = "Relationship Satisfaction"
  ) +
  scale_fill_manual(values = c("#FF9999", "#99CCFF")) +  # Customize fill colors
  scale_color_manual(values = c("#CC6666", "#6699CC")) +  # Customize jitter colors
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.title = element_blank()
  )
```
Results: Main Effects are not significant - Fail to reject the null hypothesis.
At least one Interaction between Anti-Depressant Use and Gender is significant p = .049046




```{r}
#Sexual Satisfaction

#descriptive Statistics
group_by(Selected_Dataset, GENDER, ANTIDEPRESSANT_USE) %>%
  dplyr::summarise(
    mean = mean(SEXUAL_SATISFACTION, na.rm = TRUE),
    sd = sd(SEXUAL_SATISFACTION, na.rm = TRUE)
  )

#3x2 ANOVA
Sexual_Satisfaction_Model <- aov(SEXUAL_SATISFACTION ~ GENDER * ANTIDEPRESSANT_USE, data = Selected_Dataset)

summary(Sexual_Satisfaction_Model)

# Post Hoc
TukeyHSD(Sexual_Satisfaction_Model, which = "GENDER:ANTIDEPRESSANT_USE")
```


```{r}
#Commitment

#descriptive Statistics
group_by(Selected_Dataset, GENDER, ANTIDEPRESSANT_USE ) %>%
  dplyr::summarise(
    mean = mean(COMMITMENT, na.rm = TRUE),
    sd = sd(COMMITMENT, na.rm = TRUE)
  )

#3x2 ANOVA
Commitment_Model <- aov(COMMITMENT ~ GENDER* ANTIDEPRESSANT_USE , data = Selected_Dataset)

summary(Commitment_Model)

# Post Hoc
TukeyHSD(Commitment_Model, which = "GENDER:ANTIDEPRESSANT_USE")
```

# Reliability for Relationship Satisfaction
```{r}
#Cronbach's Alpha
Alpha(Selected_Dataset, vars = c("Q17_1", "Q19", "Q20", "Q21"))
```

#Factor Analysis for Relationship Satisfaction
```{r}
#Factor Analysis
EFA(Selected_Dataset, vars = c("Q17_1", "Q19", "Q20", "Q21"), method = "pa", plot.scree = TRUE, nfactors = c("parallel"))

```


#Additional Graphs for Poster
```{r}
#Pie Chart - Genders
pct <- round(100*table(Selected_Dataset$GENDER)/length(Selected_Dataset$GENDER), 1)
table(Selected_Dataset$GENDER)

GenderLabels <- c("Men", "Women", "Other")
GenderColors <- c("white", "lightblue", "grey")
pie(table(Selected_Dataset$GENDER), 
    labels = paste(GenderLabels, sep = " ", pct, "%"),
    main = "Gender",
   col = GenderColors,
    cex = 1.7,
    cex.main = 2.5)

#Pie Chart - Sexual Orientation
pct <- round(100*table(Selected_Dataset$SEXUAL_IDENTITY)/length(Selected_Dataset$SEXUAL_IDENTITY), 1)
table(Selected_Dataset$SEXUAL_IDENTITY)

SexualLabels <- c("Other", "Heterosexual")
pie(table(Selected_Dataset$SEXUAL_IDENTITY), 
    labels = paste(SexualLabels, sep = " ", pct, "%"),
    main = "Sexual Orientation",
    cex = 1.7,
     cex.main = 2.5)


#Pie Chart - Antidepressant Use
pct <- round(100*table(Selected_Dataset$ANTIDEPRESSANT_USE)/length(Selected_Dataset$ANTIDEPRESSANT_USE), 1)
table(Selected_Dataset$ANTIDEPRESSANT_USE)

AntidepressantLabels <- c("Yes", "No")
pie(table(Selected_Dataset$ANTIDEPRESSANT_USE), 
    labels = paste(AntidepressantLabels, sep = " ", pct, "%"),
    main = "Antidepressant Use",
        cex = 1.7, 
    cex.main = 2.5)

```



